---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(knitr)
```

# Estimation of data-dependent (in)direct effects with a repeatedly measured continuous mediator and missing outcome data

## Introduction
Here we provide the R-code to reproduce the simulation study presented in our manuscript.


## Preparation
Clone the github repository, then load (and install first if necessary) the following packages
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
library(data.table)
library(sl3)
```

Source all the R functions
```{r, echo = TRUE, eval = TRUE}
for(f in list.files("R",".R$",full.names=TRUE)){source(f)}
for(f in list.files("functions",".R$",full.names=TRUE)){source(f)}
```


## Example

```{r echo = TRUE, eval = TRUE}
data <- simulateData(n=500)
```


```{r echo = TRUE, eval = TRUE}
fit <-fitLTMLE(data, t=c(1,2), L0nodes = c("L01"), Anode = "A", Cnodes = c("C1", "C2"),
               Lnodes = c("L1", "L2"), Mnodes = c("M1", "M2"), RYnode = "RY", Ynode = "Y", 
               Cmodel= list("C1 ~ A", "C2 ~ A + M1"), 
               Mmodel=list("M1 ~ A + L1", "M2 ~ M1 + A + L2"),
               gmodel=list("M1 ~ A + L1", "M2 ~ M1 + A + L2"), 
               RYmodel= "RY ~ A + M2 + L2", 
               Ymodel="Y ~ A + M2 + L2", 
               QLmodel= list("QL1 ~ L01 + A", "QL2 ~ L01 + L1 + A + M1"),
               a1 = 1, a0 = 0, n_bins = 40)
fit$est
```

Nuisance parameters can be modeled with any machine learning algorithm supported by sl3 R package

```{r echo = TRUE, eval = TRUE}
lrn_stack <- Stack$new(Lrnr_glm_fast$new(), Lrnr_mean$new(), Lrnr_bayesglm$new(), 
                       Lrnr_gam$new())
lrn_sl <- Lrnr_sl$new(learners = lrn_stack)
```

```{r echo = TRUE, eval = TRUE}
fitSL <-fitLTMLE(data, t=c(1,2), L0nodes = c("L01"), Anode = "A", Cnodes = c("C1", "C2"),
                 Lnodes = c("L1", "L2"), Mnodes = c("M1", "M2"), RYnode = "RY", Ynode = "Y", 
                 Cmodel= list("C1 ~ A", "C2 ~ A + M1"), 
                 Mmodel=list("M1 ~ A + L1", "M2 ~ M1 + A + L2"),
                 gmodel=list("M1 ~ A + L1", "M2 ~ M1 + A + L2"), 
                 RYmodel= "RY ~ A + M2 + L2", 
                 Ymodel="Y ~ A + M2 + L2", 
                 QLmodel= list("QL1 ~ L01 + A", "QL2 ~ L01 + L1 + A + M1"),
                 a1 = 1, a0 = 0, n_bins = 40,
                 Ylearner=lrn_sl, RYlearner = lrn_sl, Clearner = lrn_sl)
fitSL$est
```


## Simulation study
Our simulations studies are organized with the help of the targets package (see https://books.ropensci.org/targets/). The simulation set-up is defined in the master file ./_targets.R. The results can be assessed by the function tar_read() as shown below.

```{r}

```

