---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(knitr)
```

# Estimation of interventional effects in clinical trials with a repeatedly measured mediator

## Introduction <a name="introduction"></a>
Here we provide the R-code to reproduce the simulation study presented in our manuscript.

## Preparation <a name="preparation"></a>
Clone the github repository, then load (and install first if necessary) the following packages
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
library(data.table)
library(sl3)
```

Source all the R functions
```{r, echo = TRUE, eval = TRUE}
for(f in list.files("R",".R$",full.names=TRUE)){source(f)}
for(f in list.files("functions",".R$",full.names=TRUE)){source(f)}
```


## Example <a name="example"></a>

### Simple example <a name="simpleexample"></a>
To illustrate how the implementation works we generate a simulated data set from the data generating mechanism (ii) described in the manuscript. 
```{r echo = TRUE, eval = TRUE}
set.seed(67394)
setting_ii <- data.frame(betaL1.A=-0.25, betaM1.A=-3.50, betaL2.A=-0.15,betaM2.A=-0.70, 
                         betaL3.A=-0.10, betaM3.A=-0.25, betaL2.M=0.02, betaL3.M=0.01,
                         betaY.A=0.90, betaY.M=-0.40, betaY.L=-1.00)
data <- do.call(simulateData, c(list(n=500), setting_ii))
```

The estimates are computed using the <tt>`fitLTMLE`</tt> function 
```{r echo = TRUE, eval = TRUE}
fit <- fitLTMLE(data = data, L0nodes = c("L01", "L02"), Anode = "A", Cnodes = c("C1", "C2", "C3"),
                Lnodes = c("L1", "L2", "L3"), Mnodes = c("M1", "M2", "M3"), RYnode = "RY", Ynode = "Y", 
                Cmodel= list("C1 ~ A", "C2 ~ A + M1", "C3 ~ A + M2"), 
                Mmodel=list("M1 ~ A + L1 + L02", "M2 ~ A + M1 + L2", "M3 ~ A + M2 + L3"),
                gmodel=list("M1 ~ A + L1 + L02", "M2 ~ A + M1 + L2", "M3 ~ A + M2 + L3"),
                RYmodel= "RY ~ A + M3 + L3", Ymodel="Y ~ A + M3 + L3", 
                QLmodel= list("QL1 ~ L01 + A", "QL2 ~  A + L1 + M1", "QL3 ~ A + L2 + M2"),
                a1 = 1, a0 = 0)
```

```{r echo = TRUE, eval = TRUE}
summary(fit, type="diff", conf.int=0.95)
```

```{r echo = TRUE, eval = TRUE}
summary(fit, type="OR", conf.int=0.95)
```

### Data adaptive estimation <a name="dataadaptive"></a>
Nuisance parameters can be modeled with any machine learning algorithm supported by the [<tt>`sl3`</tt>](https://github.com/tlverse/sl3) R package. To illustrate this we define an ensemble Super Learner 
```{r echo = TRUE, eval = TRUE}
lrn_stack <- Stack$new(Lrnr_mean$new(), Lrnr_glm_fast$new(), Lrnr_bayesglm$new(), Lrnr_gam$new())
lrn_sl <- Lrnr_sl$new(
  learners = lrn_stack, 
  metalearner = Lrnr_nnls$new()
)
```

We use this learner to estimate the nuisance models $Q_Y$, $p_{R_Y}$, $Q_L$ and $p_C$ by specifying the <tt>`Ylearner`</tt>, <tt>`RYlearner`</tt>,<tt>`QLlearner`</tt>  and <tt>`Clearner`</tt> arguments as follows
```{r echo = TRUE, eval = TRUE}
fitSL <- fitLTMLE(data, L0nodes = c("L01", "L02"), Anode = "A", Cnodes = c("C1", "C2", "C3"),
                Lnodes = c("L1", "L2", "L3"), Mnodes = c("M1", "M2", "M3"), RYnode = "RY", Ynode = "Y", 
                Cmodel= list("C1 ~ A", "C2 ~ A + M1", "C3 ~ A + M2"), 
                Mmodel=list("M1 ~ A + L1 + L02", "M2 ~ A + M1 + L2", "M3 ~ A + M2 + L3"),
                gmodel=list("M1 ~ A + L1 + L02", "M2 ~ A + M1 + L2", "M3 ~ A + M2 + L3"),
                RYmodel= "RY ~ A + M3 + L3", Ymodel="Y ~ A + M3 + L3", 
                QLmodel= list("QL1 ~ L01 + A", "QL2 ~  A + L1 + M1", "QL3 ~ A + L2 + M2"),
                a1 = 1, a0 = 0, Ylearner = lrn_sl, RYlearner = lrn_sl, QLlearner = lrn_sl, 
                Clearner = lrn_sl)
```

```{r echo = TRUE, eval = TRUE}
summary(fitSL, type="diff")
```

```{r echo = TRUE, eval = TRUE}
summary(fitSL, type="OR")
```



## Simulation studies <a name="simulations"></a>

The simulations can be replicated using the script run-sim.R as shown below
```{r eval=FALSE, include=TRUE}
source("settings/simulation_settings.R")
n <- 400 # sample size
setting <- setting_ii # data generating mechanism
formula <- formula_correct_ii # model formulas 
name <- "sim1-settingII" # name of output 
source("run-sim.R")
```