---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(knitr)
```

# Estimation of interventional effects in clinical trials with a repeatedly measured mediator

## Introduction <a name="introduction"></a>
Here we provide the R-code to reproduce the simulation study presented in our manuscript.

## Preparation <a name="preparation"></a>
Clone the github repository, then load (and install first if necessary) the following packages
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
library(data.table)
library(sl3)
```

Source all the R functions
```{r, echo = TRUE, eval = TRUE}
for(f in list.files("R",".R$",full.names=TRUE)){source(f)}
for(f in list.files("functions",".R$",full.names=TRUE)){source(f)}
```


## Example <a name="example"></a>

### Simple example <a name="simpleexample"></a>
To illustrate how the implementation works we generate a simulated data set from the data generating mechanism (ii) described in the manuscript. 
```{r echo = TRUE, eval = TRUE}
set.seed(67394)
setting_ii <- list(betaL1.A=0.75, betaM1.A=0.75, betaL2.A=1.00,betaM2.A=1.00,
                        betaY.A=0.75, betaY.M=0.20, betaY.AL=0.0, betaY.L=-0.15, alphaY=-1)
data <- do.call(simulateData, c(list(n=500), setting_ii))
```

The estimates are computed using the <tt>`fitLTMLE`</tt> function 
```{r echo = TRUE, eval = TRUE}
fit <- fitLTMLE(data = data, L0nodes = c("L01"), Anode = "A", Cnodes = c("C1", "C2"),
                Lnodes = c("L1", "L2"), Mnodes = c("M1", "M2"), RYnode = "RY", Ynode = "Y", 
                Cmodel = list("C1 ~ A", "C2 ~ A + M1"), 
                Mmodel = list("M1 ~ A + L1", "M2 ~ M1 + A + L2"),
                gmodel = list("M1 ~ A + L1", "M2 ~ M1 + A + L2"), 
                RYmodel = "RY ~ A + M2 + L2", 
                Ymodel = "Y ~ A + M2 + L2", 
                QLmodel = list("QL1 ~ L01 + A", "QL2 ~ L01 + L1 + A + M1"),
                a1 = 1, a0 = 0)
```

```{r echo = TRUE, eval = TRUE}
summary(fit, type="diff", conf.int=0.95)
```

```{r echo = TRUE, eval = TRUE}
summary(fit, type="OR", conf.int=0.95)
```

### Data adaptive estimation <a name="dataadaptive"></a>
Nuisance parameters can be modeled with any machine learning algorithm supported by the [<tt>`sl3`</tt>](https://github.com/tlverse/sl3) R package. To illustrate this we define an ensemble Super Learner 
```{r echo = TRUE, eval = TRUE}
lrn_stack <- Stack$new(Lrnr_mean$new(), Lrnr_glm_fast$new(), Lrnr_bayesglm$new(), Lrnr_gam$new())
lrn_sl <- Lrnr_sl$new(
  learners = lrn_stack, 
  metalearner = Lrnr_nnls$new()
)
```

We use this learner to estimate the nuisance models $Q_Y$, $p_{R_Y}$, $Q_L$ and $p_C$ by specifying the <tt>`Ylearner`</tt>, <tt>`RYlearner`</tt>,<tt>`QLlearner`</tt>  and <tt>`Clearner`</tt> arguments as follows
```{r echo = TRUE, eval = TRUE}
fitSL <- fitLTMLE(data, L0nodes = c("L01"), Anode = "A", Cnodes = c("C1", "C2"),
                  Lnodes = c("L1", "L2"), Mnodes = c("M1", "M2"), RYnode = "RY", Ynode = "Y", 
                  Cmodel = list("C1 ~ A", "C2 ~ A + M1"), 
                  Mmodel = list("M1 ~ A + L1", "M2 ~ M1 + A + L2"),
                  gmodel = list("M1 ~ A + L1", "M2 ~ M1 + A + L2"), 
                  RYmodel = "RY ~ A + M2 + L2", 
                  Ymodel = "Y ~ A + M2 + L2", 
                  QLmodel = list("QL1 ~ L01 + A", "QL2 ~ L01 + L1 + A + M1"),
                  a1 = 1, a0 = 0, Ylearner = lrn_sl, RYlearner = lrn_sl, QLlearner=lrn_sl, Clearner = lrn_sl)
```

```{r echo = TRUE, eval = TRUE}
summary(fitSL, type="diff")
```

```{r echo = TRUE, eval = TRUE}
summary(fitSL, type="OR")
```

## Simulation studies <a name="simulations"></a>
Our simulations studies are organized with the help of the R package [<tt>`targets`</tt>](https://books.ropensci.org/targets/). The simulation set-up is defined in the master file ./_targets.R. The results can be assessed by the function <tt>`tar_read()`</tt> as shown below.

```{r eval=FALSE, include=FALSE}
sim1 <- targets::tar_read(sim1)
setDT(sim1)
sim1[, .(n=mean(n), est.sde=mean(est.sde), bias.sde = mean(est.sde-sde.true), sd.sde=sd(est.sde), se.sde=mean(se.sde), cov.sde = mean((sde.true < est.sde + qnorm(0.975)*se.sde) & (sde.true > est.sde - qnorm(0.975)*se.sde)), est.sie=mean(est.sie), bias.sie = mean(est.sie-sie.true), sd.sie=sd(est.sie), se.sie=mean(se.sie), cov.sie = mean((sie.true < est.sie + qnorm(0.975)*se.sie) & (sie.true > est.sie - qnorm(0.975)*se.sie)), est.oe=mean(est.oe), bias.oe = mean(est.oe-oe.true), sd.oe=sd(est.oe), se.oe=mean(se.oe), cov.oe = mean((oe.true < est.oe + qnorm(0.975)*se.oe) & (oe.true > est.oe - qnorm(0.975)*se.oe)), est.pm=mean(est.pm)), by=tar_group]
```

```{r eval=FALSE, include=FALSE}
sim1[, .(n=mean(n), est.sde=mean(est.ORsde), bias.sde = mean(est.ORsde-sde.OR.true), sd.sde=sd(est.ORsde), se.sde=mean(se.ORsde), cov.sde = mean((sde.OR.true < est.ORsde + qnorm(0.975)*se.ORsde) & (sde.OR.true > est.ORsde - qnorm(0.975)*se.ORsde)), est.sie=mean(est.ORsie), bias.sie = mean(est.ORsie-sie.OR.true), sd.sie=sd(est.ORsie), se.sie=mean(se.ORsie), cov.sie = mean((sie.OR.true < est.ORsie + qnorm(0.975)*se.ORsie) & (sie.OR.true > est.ORsie - qnorm(0.975)*se.ORsie)), est.oe=mean(est.ORoe), bias.oe = mean(est.ORoe-oe.OR.true), sd.oe=sd(est.ORoe), se.oe=mean(se.ORoe), cov.oe = mean((oe.OR.true < est.ORoe + qnorm(0.975)*se.ORoe) & (oe.OR.true > est.ORoe - qnorm(0.975)*se.ORoe))), by=tar_group]
```

```{r eval=FALSE, include=FALSE}
sim2 <- targets::tar_read(sim2)
setDT(sim2)
sim2[, .(est.sde=mean(est.sde), bias.sde = mean(est.sde-sde.true), sd.sde=sd(est.sde), se.sde=mean(se.sde), cov.sde = mean((sde.true < est.sde + qnorm(0.975)*se.sde) & (sde.true > est.sde - qnorm(0.975)*se.sde)), est.sie=mean(est.sie), bias.sie = mean(est.sie-sie.true), sd.sie=sd(est.sie), se.sie=mean(se.sie), cov.sie = mean((sie.true < est.sie + qnorm(0.975)*se.sie) & (sie.true > est.sie - qnorm(0.975)*se.sie)), est.oe=mean(est.oe), bias.oe = mean(est.oe-oe.true), sd.oe=sd(est.oe), se.oe=mean(se.oe), cov.oe = mean((oe.true < est.oe + qnorm(0.975)*se.oe) & (oe.true > est.oe - qnorm(0.975)*se.oe))), by=.(n, mis)]
```

```{r eval=FALSE, include=FALSE}
sim2[, .(n=mean(n), est.sde=mean(est.ORsde), bias.sde = mean(est.ORsde-sde.OR.true), sd.sde=sd(est.ORsde), se.sde=mean(se.ORsde), cov.sde = mean((sde.OR.true < est.ORsde + qnorm(0.975)*se.ORsde) & (sde.OR.true > est.ORsde - qnorm(0.975)*se.ORsde)), est.sie=mean(est.ORsie), bias.sie = mean(est.ORsie-sie.OR.true), sd.sie=sd(est.ORsie), se.sie=mean(se.ORsie), cov.sie = mean((sie.OR.true < est.ORsie + qnorm(0.975)*se.ORsie) & (sie.OR.true > est.ORsie - qnorm(0.975)*se.ORsie)), est.oe=mean(est.ORoe), bias.oe = mean(est.ORoe-oe.OR.true), sd.oe=sd(est.ORoe), se.oe=mean(se.ORoe), cov.oe = mean((oe.OR.true < est.ORoe + qnorm(0.975)*se.ORoe) & (oe.OR.true > est.ORoe - qnorm(0.975)*se.ORoe))),  by=.(n, mis)]
```

```{r eval=FALSE, include=FALSE}
sim3 <- targets::tar_read(sim3)
setDT(sim3)
sim3[, .(n=mean(n),est.sde=mean(est.sde), bias.sde = mean(est.sde-sde.true), sd.sde=sd(est.sde), se.sde=mean(se.sde), cov.sde = mean((sde.true < est.sde + qnorm(0.975)*se.sde) & (sde.true > est.sde - qnorm(0.975)*se.sde)), est.sie=mean(est.sie), bias.sie = mean(est.sie-sie.true), sd.sie=sd(est.sie), se.sie=mean(se.sie), cov.sie = mean((sie.true < est.sie + qnorm(0.975)*se.sie) & (sie.true > est.sie - qnorm(0.975)*se.sie)), est.oe=mean(est.oe), bias.oe = mean(est.oe-oe.true), sd.oe=sd(est.oe), se.oe=mean(se.oe), cov.oe = mean((oe.true < est.oe + qnorm(0.975)*se.oe) & (oe.true > est.oe - qnorm(0.975)*se.oe))), by=tar_group]
```

```{r eval=FALSE, include=FALSE}
sim3[, .(n=mean(n), est.sde=mean(est.ORsde), bias.sde = mean(est.ORsde-sde.OR.true), sd.sde=sd(est.ORsde), se.sde=mean(se.ORsde), cov.sde = mean((sde.OR.true < est.ORsde + qnorm(0.975)*se.ORsde) & (sde.OR.true > est.ORsde - qnorm(0.975)*se.ORsde)), est.sie=mean(est.ORsie), bias.sie = mean(est.ORsie-sie.OR.true), sd.sie=sd(est.ORsie), se.sie=mean(se.ORsie), cov.sie = mean((sie.OR.true < est.ORsie + qnorm(0.975)*se.ORsie) & (sie.OR.true > est.ORsie - qnorm(0.975)*se.ORsie)), est.oe=mean(est.ORoe), bias.oe = mean(est.ORoe-oe.OR.true), sd.oe=sd(est.ORoe), se.oe=mean(se.ORoe), cov.oe = mean((oe.OR.true < est.ORoe + qnorm(0.975)*se.ORoe) & (oe.OR.true > est.ORoe - qnorm(0.975)*se.ORoe))), by=tar_group]
```

```{r eval=FALSE, include=FALSE}
tableD1 <- targets::tar_read(sim_bins)
setDT(tableD1)
tableD1[, .(est.psi00=mean(est.psi00), bias.psi00 = mean(est.psi00-psi00.true), sd.psi00=sd(est.psi00), se.psi00=mean(se.psi00), cov.psi00 = mean((psi00.true < est.psi00 + qnorm(0.975)*se.psi00) & (psi00.true > est.psi00 - qnorm(0.975)*se.psi00)), est.psi11=mean(est.psi11), bias.psi11 = mean(est.psi11-psi11.true), sd.psi11=sd(est.psi11), se.psi11=mean(se.psi11), cov.psi11 = mean((psi11.true < est.psi11 + qnorm(0.975)*se.psi11) & (psi11.true > est.psi11 - qnorm(0.975)*se.psi11)), est.psi10=mean(est.psi10), bias.psi10 = mean(est.psi10-psi10.true), sd.psi10=sd(est.psi10), se.psi10=mean(se.psi10), cov.psi10 = mean((psi10.true < est.psi10 + qnorm(0.975)*se.psi10) & (psi10.true > est.psi10 - qnorm(0.975)*se.psi10))), by=.(n, n_bins)]
```

```{r eval=FALSE, include=FALSE}
tableD1[, .(est.sde=mean(est.sde), bias.sde = mean(est.sde-sde.true), sd.sde=sd(est.sde), se.sde=mean(se.sde), cov.sde = mean((sde.true < est.sde + qnorm(0.975)*se.sde) & (sde.true > est.sde - qnorm(0.975)*se.sde)), est.sie=mean(est.sie), bias.sie = mean(est.sie-sie.true), sd.sie=sd(est.sie), se.sie=mean(se.sie), cov.sie = mean((sie.true < est.sie + qnorm(0.975)*se.sie) & (sie.true > est.sie - qnorm(0.975)*se.sie)), est.oe=mean(est.oe), bias.oe = mean(est.oe-oe.true), sd.oe=sd(est.oe), se.oe=mean(se.oe), cov.oe = mean((oe.true < est.oe + qnorm(0.975)*se.oe) & (oe.true > est.oe - qnorm(0.975)*se.oe))), by=.(n, n_bins)]
```

```{r eval=FALSE, include=FALSE}
tableD1[, .(n=mean(n), est.sde=mean(est.ORsde), bias.sde = mean(est.ORsde-sde.OR.true), sd.sde=sd(est.ORsde), se.sde=mean(se.ORsde), cov.sde = mean((sde.OR.true < est.ORsde + qnorm(0.975)*se.ORsde) & (sde.OR.true > est.ORsde - qnorm(0.975)*se.ORsde)), est.sie=mean(est.ORsie), bias.sie = mean(est.ORsie-sie.OR.true), sd.sie=sd(est.ORsie), se.sie=mean(se.ORsie), cov.sie = mean((sie.OR.true < est.ORsie + qnorm(0.975)*se.ORsie) & (sie.OR.true > est.ORsie - qnorm(0.975)*se.ORsie)), est.oe=mean(est.ORoe), bias.oe = mean(est.ORoe-oe.OR.true), sd.oe=sd(est.ORoe), se.oe=mean(se.ORoe), cov.oe = mean((oe.OR.true < est.ORoe + qnorm(0.975)*se.ORoe) & (oe.OR.true > est.ORoe - qnorm(0.975)*se.ORoe))), by=.(n, n_bins)]
```
